---
title: "003_note_data_tidying"
author: "Mariia Koliada"
date: "28 02 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#two primary packages in R that will help you reshape your data: tidyr and reshape2
#tidyr
#Within tidyr, there are two functions to help you reshape your data.
#gather(): go from wide data to long data
#spread(): go from long data to wide data

tidyr package is installed and loaded into your RStudio session.

```{r}
install.packages("tidyr")
library(tidyr)
```

## use gather() to reshape from wide to long
```{r}
gathered <- gather(airquality)
```

## take a look at first few rows of long data
head(gathered)
```{r}
head(gathered)
```
## to rename the column names that gather provides,
## change key and value to what you want those column names to be
gathered <- gather(airquality, key="variable", value="value")
```{r}
gathered <- gather(airquality, key ="variable", value="value")
```


## take a look at first few rows of long data
head(gathered)

```{r}
## in gather(), after key and value, you can specify which variables 
## you want included in the long format
## it will leave the other variables as is
gathered <- gather(airquality, key="variable", value="value", ozone, solar.r, wind, temp)

## take a look at first few rows of long data
head(gathered)
```

# View the first 6 rows
head(iris)

# View the first 15 rows
head(iris, n = 15)

# View the last 6 rows
tail(iris)

# View the last 10 rows
tail(iris, n = 10)


